---
title: "EDA Amjad"
output: html_document
date: "2022-10-06"
---
#Data loading and structure
```{r}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyr)

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
bike <- read.csv("bo.csv")
str(bike)
```


```{r}
head(bike, n=4)
```

#Data preparing 
```{r}
# change rideable_type/ member_casual to factor 
#there three numbers under the factor rideable_type needs to look into!
bike$rideable_type <- factor(bike$rideable_type)
bike$member_casual <- factor(bike$member_casual)

# split started_at/ ended_at to date column and time column
# change start_date/end_date to date type
bike$start_time <- format(as.POSIXct(bike$started_at), format = "%H:%M:%S") 
bike$end_time <- format(as.POSIXct(bike$ended_at), format = "%H:%M:%S") 
bike$start_date <- as.Date(bike$started_at)
bike$end_date <- as.Date(bike$ended_at)

# change time column to time column

# Check the structure again
str(bike)
```

#Data analysis & visualazation
**How many users of each membership type we have?**
```{r}

```
**What is the most frequent bike type used?**
```{r}

```
**Which day of the week the serves is used more?**
```{r}

```
**When is the highest-lowest time of use of the day?**
```{r}

```
**Is there a correlation between time and day of usage?**
```{r}

```
**What is the average trip distance?**
```{r}

```
**where are the most used stations**
#heat map
```{r}

```