---
title: "EDS"
output: html_document
date: "2022-11-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(purrr)   #to join all files
library(lubridate)# for duration of time
library(ModelMetrics)
library(pROC)
library(dplyr)
library(tidyr)
library(caret)
library(ggplot2)
```


#joining data for 6 months in one sheet
```{r}
#bind all the files in this path together
file.list <- list.files(pattern='*.csv')

# need an id-column with the file-names
file.list <- setNames(file.list, file.list)

#want only sheet 4, skip 1 row and name id column month
bike <-map_df(file.list, function(x) read.csv(x))

str(bike)
```

```{r}
bike%>%group_by(rideable_type)%>%
  summarise(count = n())

#Remove docked bikes
bike <-subset(bike, rideable_type != "docked_bike")

bike%>%group_by(rideable_type)%>%
  summarise(count = n())

#70% is unbalanced data set and this cut is more than that
```
```{r}

dim(bike)

quartiles <- quantile(bike$Sepal.Width, probs=c(.25, .75), na.rm = FALSE)
IQR <- IQR(bike$Sepal.Width)
 
Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 
 
data_no_outlier <- subset(bike, bike$Sepal.Width > Lower & bike$Sepal.Width < Upper)
 
dim(data_no_outlier)

data_no_outlier
```

#Preparing columns for modeling
```{r}

bike$rideable_type <- factor(as.integer(factor(bike$rideable_type)))
bike$start_station_id <- factor(bike$start_station_id)
bike$end_station_id <- factor(bike$end_station_id)
# make it to casual = 0, member= 1 
bike$member_casual<-factor(ifelse(bike$member_casual=='casual',0,1))

# split started_at/ ended_at to date column and time column
# change start_date/end_date to date type
bike$start_time <- format(as.POSIXct(bike$started_at), format = "%H:%M")
bike$end_time <- format(as.POSIXct(bike$ended_at), format = "%H:%M") 

# covert date to date type to get the time differnce
bike$Startedat = as.POSIXct(bike$started_at, format = "%Y-%m-%d %H:%M:%S")
bike$enddat = as.POSIXct(bike$ended_at, format = "%Y-%m-%d %H:%M:%S")
#bike$Total_Time <- round(as.numeric(difftime(bike$enddat, bike$Startedat, units = "mins")),2)

bike$Total_Time <- abs(as.numeric(difftime(bike$enddat, bike$Startedat, units = "mins")))

# take only the day from the date
bike$start_date <- as.Date(bike$started_at)
bike$end_date <- as.Date(bike$ended_at)
bike$days <- factor(format(bike$start_date, format = "%a"))

bike

```


#make the time as a duration of mints
```{r}
res <- hm(bike$start_time)# format to 'hours:minutes'
bike$startingm <- hour(res)*60 + minute(res)

res <- hm(bike$end_time)# format to 'hours:minutes'
bike$endingm <- hour(res)*60 + minute(res)

#one hot encoding of days
#bike <-bike %>% mutate(value = 1)  %>% spread(days, value,  fill = 0 )
```


## Drop non-needed columns
```{r}
#selecting the columns we want
want <- c('member_casual','rideable_type', 'Total_Time', 'start_station_id', 'end_station_id', 'startingm','endingm','days')

df2 <- bike[,(names(bike) %in% want)]

#Delete the na 
df<-na.omit(df2)
df %>% drop_na()

#make member_casual to be the first column
colnames(df)
df <- df %>%
  select('rideable_type', everything())

```

# Check if we have unbalanced data

```{r}
df_rt%>%group_by(rideable_type)%>%
  summarise(count = n())

df_rt %>% count(rideable_type)
ggplot(df_rt, aes(rideable_type, fill = rideable_type))+
  geom_bar()+
  scale_fill_brewer(palette = "BuPu")+
  guides(fill="none")+
  labs(title = "User bike types", x= "bike types")+
  theme_classic()
```

#Test and train split
```{r}
#scaleddf <- as.data.frame(scale(df[2:13], center = TRUE, scale = TRUE))
set.seed(333)
#df_sample_rt <- sample(2, nrow(df_rt), replace=TRUE, prob=c(0.80, 0.20))


#Rt_train_X <- df_rt[df_sample_rt==1, 2:7]
#Rt_test_X <- df_rt[df_sample_rt==2, 2:7]


#Rt_train_y <- df_rt[df_sample_rt==1, 1]
#Rt_test_y <- df_rt[df_sample_rt==2, 1]

df_sample <- sample(2, nrow(df), replace=TRUE, prob=c(0.75, 0.25))


train_X <- df[df_sample==1, 2:8]
test_X <- df[df_sample==2, 2:8]


train_y <- df[df_sample==1, 1]
test_y <- df[df_sample==2, 1]

str(train_X)

```

# Logistic model
```{r}

Bike_Type_Logit <- glm(train_y ~ start_station_id+ end_station_id+Total_Time+member_casual+ startingm*days + endtingm , data = train_X, family = "binomial")
summary(customerLogit)

```


# evaluation
```{r}
prob=predict(Bike_Type_Logit, type = "response")
e <- roc(train_y~prob)
auc(e) # area-under-curve 
plot(e)

```

#Check the accurcy for the test
```{r}
proc=predict(Bike_Type_Logit, type = "response", newdata = test_X)
test_X$proc=proc
roc <- roc(test_y~proc)
print(roc)

```

# See how well the model is doing by running suveral tests:
## Deviance tests



## McFadden
```{r}
library(pscl)
pR2(Bike_Type_Logit)
```


### Akaike Information Criterion (AIC)
```{r}
AIC(Bike_Type_Logit)

```
Lower AIC is better but I don't know if this is lower?


```{r}
# cchange the 
pred_1 = predict(Bike_Type_Logit, test_X, type="response")
paste0("Accuracy = ",mean(Bike_Type_Logit$rideable_type == pred_1))

bt$prob = as.numeric(unlist(as.data.frame(pred_1)[2]))
bt$rideable_type = as.numeric(as.character(bt$rideable_type))
bt$prob = round(bt$prob)
h <- roc(rideable_type~prob, data=Rt_test_X)
print(h)
plot(h,main="ROC Curve for Random Forest",col=2,lwd=2)


```

# Make prediction
## see 0.5 cut
```{r}

logitpredict.5 <- predict(Bike_Type_Logit, newdata = test_X, type = "response") > 0.55

## I don't know if members are 1 or 0 
defult <- ifelse(logitpredict.5 =="TRUE", 1, 0)
crossTable = table(defult,test_y)
t <- as.data.frame(crossTable)
crossTable
```

### plot confution metrix
```{r}
ggplot(data =  t, mapping = aes(x = defult, y = test_y)) +
  geom_tile(aes(fill = Freq), colour = "white") +
  geom_text(aes(label = sprintf("%1.0f", Freq)), vjust = 1) +
  scale_fill_gradient(low = "#98c1d9", high = "#8f2d56") +
  theme_bw() + 
  theme(legend.position = "none") +
  labs(title = "Confution Matrix with Cutoff at 0.05")+
  xlab( "predected")+
  ylab( "original")
```

###The accuracy of the model of 0.5 cutoff :
```{r}
format(100*(crossTable[1,1]+crossTable[2,2])/sum(crossTable), digits=4)
```

## see 0.4 cut
```{r}

logitpredict.4 <- predict(customerLogit, newdata = test_X, type = "response") > 0.6
defult2 <- ifelse(logitpredict.4 =="TRUE", 1, 0)
crossTable4 = table(defult2,test_y)
d <- as.data.frame(crossTable4)
crossTable4
```


### The accuracy of the model of 0.4 cutoff :
```{r}
format(100*(crossTable4[1,1]+crossTable4[2,2])/sum(crossTable4), digits=4)
```
```{r}

ggplot(data = d , mapping = aes(x = defult2, y = test_y)) +
  geom_tile(aes(fill = Freq), colour = "white") +
  geom_text(aes(label = sprintf("%1.0f", Freq)), vjust = 1) +
  scale_fill_gradient(low = "#98c1d9", high = "#8f2d56") +
  theme_bw() +
  theme(legend.position = "none")+
  labs(title = "Confution Matrix with Cutoff at 0.04")+
  xlab( "predected")+
  ylab( "original")

```
